---
title: "Kaggle House Prices Competition"
output: html_notebook
---


Load the data
```{r}
dat <- read.csv("train.csv")
str(dat)
```

Wow, there are a lot of featutres! It's going to be challenging to sort through these and get a good overview of what's going on..

First let's look at distribution of sale price
```{r}
library(ggplot2)
ggplot(dat,aes(x=SalePrice))+
        geom_histogram()

ggplot(dat,aes(x=1,y=SalePrice))+
        geom_boxplot()
```


Let's see if there is a trend over the past few years (might be influenced by overall US housing market). There doesn't seem to be much of a trend vs year.
```{r}
ggplot(dat, aes(YrSold,SalePrice))+
        geom_boxplot(aes(group=YrSold))
```


My first intution is that the price will be highly dependent on the size of the house, size of lot, and number of bedrooms . Let's check if that is true:
```{r}

# vs #bedrm above grade: linear trend from 2-4 bedrooms
ggplot(dat, aes(BedroomAbvGr,SalePrice))+
        geom_boxplot(aes(group=BedroomAbvGr))

# price vs lot size
library(dplyr)
ggplot(filter(dat,LotArea<50000),aes(LotArea,SalePrice))+
        geom_point()+
        geom_smooth(method="lm")

```

Let's looks at house prices by neighborhood. There are definitelya couple of neighborhoods which are have higher sale prices (for example NridgHt, NoRidge).
```{r}
ggplot(dat, aes(Neighborhood,SalePrice))+
        geom_boxplot(aes(group=Neighborhood))+
        coord_flip()
```


Ok there are too many variables to look at each one individually. Let's try computing the correlation of each variable with sale price to see which ones might be more important:
```{r}
# figure out which variables are numeric
is_num <- sapply(dat,is.numeric)
id_num <- which(is_num)
names_num=names(is_num)[id_num]

# compute correlation of each numeric variable w/ SalesPrice
allc <- vector(length(id_num),mode="numeric")
for (i in seq_along(id_num)){
        allc[i] <- cor(dat[,id_num[i]] , dat$SalePrice)
}

# put the results in a data frame so we can sort easily
cor_df <- data_frame(var=names_num,cor=allc)
cor_df %>%
        mutate(abs_cor=abs(cor)) %>%
        arrange(desc(abs_cor))
```

```{r}
ggplot(dat,aes(OverallQual,SalePrice))+
        geom_point()

ggplot(dat,aes(GrLivArea,SalePrice))+
        geom_point()

ggplot(dat,aes(factor(GarageCars),SalePrice))+
        geom_boxplot()
```

